Site location study for a high-mountain millimeter observatory in the Colombian Andes I: In situ data

### 2017-03-02 pasp/paperclimate.tex
intro

Astronomical observations in the millimeter and sub-millimeter wavelength range require that atmospheric effects affecting absorption at these wavelengths are kept to a minimum. The main factor contributing to the atmospheric opacity is water vapor, which very efficiently absorbs light in the THz range due to BLAHBLAHBLAH. In order to characterize a site according to its atmospheric transparency to THz radiation, it is necessary to retrieve the precipitable water vapor profile either via remote sensing techniques such as microwave radiometry, satellite measurements, or indirectly via models of in situ climatological measurements or GPS-delay studies. \\

Most of the precipitable water vapor that can potentially affect the path of a THz photon hoping to get through the atmosphere, actually exists in the troposphere. The lower the elevation of a potential site, the longer the path becomes for that photon, increasing its chances of being absorbed by a water vapor molecule. For this reason, the geographic characterization of the quality of a site is given in terms of the local water vapor column above the potential site. This water vapor column is usually expressed as the amount of precipitable water vapor, given in mm.\\

>>>>

Astronomical observations in the millimeter and sub-millimeter wavelength range require that atmospheric effects affecting absorption at these wavelengths are kept to a minimum \cite{chajpwv,radford2016}. The main factor contributing to the atmospheric opacity is water vapor, which very efficiently absorbs light in the THz range due to BLAHBLAHBLAH. In order to characterize a site according to its atmospheric transparency to THz radiation, it is necessary to retrieve the precipitable water vapor profile either via remote sensing techniques such as microwave radiometry \cite{radiometer}, satellite measurements, or indirectly via models of in situ climatological measurements or GPS-delay studies. \\

Most of the precipitable water vapor that can potentially affect the path of a THz photon hoping to get through the atmosphere actually exists in the troposphere. The lower the elevation of a potential site, the longer the path becomes for that photon, increasing its chances of being absorbed by a water vapor molecule. For this reason, the geographic characterization of the quality of a site is given in terms of the local water vapor column above the potential site. This water vapor column is usually expressed as the amount of precipitable water vapor, given in mm.\\
